---
title: "P8105 Midterm Project"
author: "Junxian Chen (jc5314)"
date: "10/18/2019"
output: github_document
---

# Introduction:

# Problem 1

Load the R packages:

```{r, include=FALSE}
library(readxl)
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_gray(base_size = 10) + theme(legend.position = "bottom"))
```

Load, clean and tidy the dataset (the raw dataset and the dataset after cleaning were seperated so that we can go back and check the raw data when possible mistakes were found):

```{r}
raw_data = 
  read_excel("./data/p8105_mtp_data.xlsx", skip = 8)
```


```{r}
mtp_data =   
  raw_data %>% 
  janitor::clean_names() %>% 
  replace_na(list(eop_size_mm = 0, eop_shape = 0)) %>% 
  mutate(
    sex = as.character(sex),
    age = as.integer(age),
    age_group = factor(age_group, levels = c('1', '2', '3', '4', '5', '6', '7', '8')),
    eop_size = factor(eop_size, levels = c('0', '1', '2', '3', '4', '5')),
    eop_visibility_classification = factor(eop_visibility_classification, levels = c('0', '1', '2')),
    eop_shape = as.character(eop_shape),
    fhp_category = factor(fhp_category, levels = c('0', '1', '2', '3', '4', '5', '6', '7'))
  )
```


```{r}
mtp_data %>% 
  mutate(
    age_group = case_when(
      age_group == '1' ~ 'less than 18',
      age_group == '2' ~ '18-30',
      age_group == '3' ~ '31-40',
      age_group == '4' ~ '41-50',
      age_group == '5' ~ '51-60',
      age_group %in% c('6', '7', '8') ~ 'larger than 60'
    )
  ) %>% 
  group_by(sex, age_group) %>% 
  summarize(n = n()) %>%
  pivot_wider(
    names_from = sex,
    values_from = n
  ) %>% 
  knitr::kable(
    col.names = c('Age', 'Female', 'Male'),
  )
```

Dataset description:


Issues in the dataset:

(1) One mistake in the age group '1'.

```{r}
mtp_data %>% 
  ggplot(aes(x = age_group, y = age, color = age_group)) +
  geom_point() +
  geom_hline(yintercept = c(18, 31, 41, 51, 61), color = 'red', linetype = 'dotted') + 
  xlab('Age Group') +
  ylab('Age (years)') +
  scale_color_hue(name = "age group")
```

(2)

```{r}
mtp_data %>% 
  ggplot(aes(x = eop_size, y = eop_size_mm, color = eop_size)) +
  geom_point() +
  geom_hline(yintercept = c(5, 10, 15, 20, 25), color = 'red', linetype = 'dotted') + 
  xlab('EOP Size Group') +
  ylab('EOP Size (mm)') +
  scale_color_hue(name = "age group")
```

Lots of mistakes in the classification in EOP size. Also, the 'NA' in the plots indicates a mistake value in EOP size grouping:

```{r}
row_na = which(is.na(pull(mtp_data, eop_size)))
raw_data[row_na,] %>% 
  knitr::kable()
```

```{r}
mtp_data %>% 
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = eop_visibility_classification)) +
  geom_point() +
  geom_hline(yintercept = c(0, 1, 5), color = 'red', linetype = 'dotted') + 
  xlab('EOP Size Group') +
  ylab('EOP Size (mm)') +
  scale_color_hue(name = "EOP size group")
```

Lots of mistakes in the classification in EOP visibility.

(3) 

```{r}
mtp_data %>% 
  ggplot(aes(x = fhp_category, y = fhp_size_mm, color = fhp_category)) +
  geom_point() +
  geom_hline(yintercept = c(0, 10, 20, 30, 40, 50, 60, 70), color = 'red', linetype = 'dotted') + 
  xlab('FHP Size Group') +
  ylab('FHP Size (mm)') +
  scale_color_hue(name = "FHP size group")
```

```{r}
row_na = which(is.na(pull(mtp_data, fhp_category)))
raw_data[row_na,] %>% 
  knitr::kable()
```

# Problem 2

```{r}
plot_df = 
  mtp_data %>% 
  mutate(
    age_group = factor(age_group, labels = c('1', '2', '3', '4', '5', '6', '6', '6')),
    droplevels(age_group)) %>% 
  filter(age_group != '1')
```

```{r}
FHP_plot =
  plot_df %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot() +
  stat_summary(
    fun.y = mean, aes(colour = "Mean"), 
    geom = "point", 
    shape = 18, 
    size = 3,
    position = position_dodge(width = 0.75)) +
  stat_summary(
    fun.data = mean_se, 
    geom = "errorbar",
    width = 0.08,
    position = position_dodge(width = 0.75)) +
  scale_fill_hue(
    name = "Sex", 
    labels = c("Females", "Males")) +
  scale_colour_manual(
    values = c("Mean" = "darkred"), 
    labs(colour = '')) +
  xlab('Age Group (years)') +
  ylab('FHP Size (mm)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
```


```{r}
EEOP_plot = 
 plot_df %>% 
  mutate(
    eeop = case_when(
      eop_size_mm > 10 ~ 'y',
      eop_size_mm <= 10 ~ 'n'
    )
  ) %>% 
  group_by(age_group, sex) %>% 
  count(eeop) %>%
  mutate(rate = prop.table(n)) %>% 
  filter(eeop == 'y') %>% 
  ggplot(aes(y = rate, x = age_group)) +
  geom_line(aes(group = sex, color = sex)) +
  geom_point() +
  scale_color_hue(
    name = "Sex", 
    labels = c("Females", "Males")) +
  xlab('Age Group (years)') +
  ylab('EEOP Rate (%)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
```


```{r}
# create a two-panel figure and show the plot

FHP_plot / EEOP_plot
```


```{r}
fhp_eop_plot = 
  plot_df %>% 
  drop_na(fhp_size_mm) %>% 
  filter(eop_size_mm != 0) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm, color = sex)) +
  geom_point() +
  geom_smooth(
    method = "lm", 
    color = 'black',
    se = FALSE) +
  facet_grid(
    sex ~ age_group, 
    labeller = as_labeller(c('0' = 'Female', '1' = 'Male', '2' = '18-29 years', 
                             '3' = "30's", '4' = "40's", '5' = "50's", '6' = ">60's"))) +
  xlab('FHP Size (mm)') +
  ylab('EOP Size (mm)') +
  scale_color_hue(
    name = "Sex", 
    labels = c("Females", "Males"))

# show the plot

fhp_eop_plot
```




