---
title: "P8105 Midterm Project"
author: "Junxian Chen (jc5314)"
date: "10/18/2019"
output: github_document
---

```{r, include = FALSE}
library(readxl)
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_gray(base_size = 10) + theme(legend.position = "bottom"))
```


## Introduction:

An article described relationship between posture and enlarged protuberances. The researchers analyzed data from individuals with different ages and sex, and concluded that prominent exostosis were more prevalent in young adults than older groups. However, skepticism arose about this study and a published author correction indicated errors existed in the original work. In this project, dataset from this study was used to test whether issues exist and reproduced results are consistent with that reported in the article.


## 1. Dataset

### 1.1 Description

```{r}
raw_data = 
  read_excel("./data/p8105_mtp_data.xlsx", skip = 8)
```

```{r}
mtp_data =   
  raw_data %>% 
  janitor::clean_names() %>% 
  replace_na(list(eop_size_mm = 0, eop_shape = 0)) %>% 
  mutate(
    sex = as.character(sex),
    age = as.integer(age),
    age_group = case_when(
      age_group %in% c('6', '7', '8') ~ '6+', 
      TRUE ~ age_group ),
    age_group = factor(age_group, levels = c('1', '2', '3', '4', '5', '6+')),
    eop_size = factor(eop_size, levels = c('0', '1', '2', '3', '4', '5')),
    eop_visibility_classification = factor(eop_visibility_classification, levels = c('0', '1', '2')),
    eop_shape = as.character(eop_shape),
    fhp_category = factor(fhp_category, levels = c('0', '1', '2', '3', '4', '5', '6', '7'))
  )

str(mtp_data)
```

In the data cleaning, missing values were filled based on given information, variable types were modified, and orders were given to factor variables. Individuals older than 60 were all assigned as '6+' under the `age_group` variables to facilitate further analyses. Because the raw data would be used later to illustrate mistakes, the process of importing and cleaning were separated here.

The resulting dataset `mtp_data` contains `r nrow(mtp_data)` observations (i.e. the number of participants) and `r ncol(mtp_data)` variables. Key variables: The average **age** was `r mean(pull(mtp_data, age))`. The average **EOP size** was `r mean(pull(mtp_data, eop_size_mm))` mm. The average **FHP size** was `r mean(pull(mtp_data, fhp_size_mm), na.rm = TRUE)` mm. `NA` exist in the `fhp_size_mm` and `eop_shape`. Since no information was provided, the `NA`'s here were left blank.

A table below shows the age and gender distribution.

```{r}
mtp_data %>% 
  mutate(
    age_group = case_when(
      age_group == '1' ~ 'less than 18',
      age_group == '2' ~ '18-30',
      age_group == '3' ~ '31-40',
      age_group == '4' ~ '41-50',
      age_group == '5' ~ '51-60',
      age_group == '6+' ~ 'larger than 60'
    )
  ) %>% 
  group_by(sex, age_group) %>% 
  summarize(n = n()) %>%
  pivot_wider(
    names_from = sex,
    values_from = n
  ) %>% 
  knitr::kable(
    col.names = c('Age', 'Female', 'Male'),
  )
```

### 1.2 Issues:

```{r}
mtp_data %>% 
  ggplot(aes(x = age_group, y = age, color = age_group)) +
  geom_point() +
  geom_hline(yintercept = c(18, 31, 41, 51, 61), color = 'red', linetype = 'dotted') + 
  xlab('Age Group') +
  ylab('Age (years)') +
  scale_color_hue(name = "age group") +
  scale_x_discrete(
    labels = c("<18's", "18-29 years","30's", "40's", "50's", ">60's")) +
  scale_color_hue(name = "age group",
                  labels = c("<18's", "18-29 years","30's", "40's", "50's", ">60's"))
```

* An outlier in '<18's' group indicates a person older than 18 was mistakenly labeled into this group. 

```{r}
mtp_data %>% 
  ggplot(aes(x = eop_size, y = eop_size_mm, color = eop_size)) +
  geom_point() +
  geom_hline(yintercept = c(5, 10, 15, 20, 25), color = 'red', linetype = 'dotted') + 
  xlab('EOP Size Group') +
  ylab('EOP Size (mm)') +
  scale_color_hue(name = "eop size group")
```

* Misclassifications in EOP size group, especially in group 1, 3, 4, 5. 

```{r}
row_na = which(is.na(pull(mtp_data, eop_size)))
raw_data[row_na,] %>% 
  knitr::kable()
```

The `NA` in x-axis is due to an abnormal value '14.6' in 'EOP size' in raw dataset which supposed to be 0 ~ 5.

```{r}
mtp_data %>% 
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = eop_visibility_classification)) +
  geom_point() +
  geom_hline(yintercept = c(0, 5), color = 'red', linetype = 'dotted') + 
  xlab('EOP Visibility Classification') +
  ylab('EOP Size (mm)') +
  scale_color_hue(name = "EOP visibility classification")
```

* Misclassifications exist in all three groups in EOP visibility classification.

```{r}
mtp_data %>% 
  ggplot(aes(x = fhp_category, y = fhp_size_mm, color = fhp_category)) +
  geom_point() +
  geom_hline(yintercept = c(0, 10, 20, 30, 40, 50, 60, 70), color = 'red', linetype = 'dotted') + 
  xlab('FHP Size Group') +
  ylab('FHP Size (mm)') +
  scale_color_hue(name = "FHP size group")
```

* Misclassifications exist in FHP size group, especially in group 1, 2, 4. 

```{r}
row_na = which(is.na(pull(mtp_data, fhp_category)))
raw_data[row_na,] %>% 
  knitr::kable()
```

The `NA` in the x-axis is due to an abnormal value '30.8' in 'FHP Category' in raw dataset which supposed to be 0 ~ 7.


## 2. Visualization Improvement

```{r}
fhp_plot =
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot() +
  stat_summary(
    fun.y = mean, aes(colour = "Mean"), 
    geom = "point", 
    shape = 18, 
    size = 3,
    position = position_dodge(width = 0.75)) +
  stat_summary(
    fun.data = mean_se, 
    geom = "errorbar",
    width = 0.15,
    position = position_dodge(width = 0.75)) +
  scale_fill_hue(
    name = "Sex", 
    labels = c("Females", "Males")) +
  scale_colour_manual(
    values = c("Mean" = "darkred"), 
    labs(colour = '')) +
  xlab('Age Group (years)') +
  ylab('FHP Size (mm)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
```


```{r}
eeop_plot = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  mutate(
    eeop = case_when(
      eop_size_mm > 10 ~ 'y',
      eop_size_mm <= 10 ~ 'n'
    )
  ) %>% 
  group_by(age_group, sex) %>% 
  count(eeop) %>%
  mutate(rate = prop.table(n)) %>% 
  filter(eeop == 'y') %>% 
  ggplot(aes(y = rate, x = age_group)) +
  geom_line(aes(group = sex, color = sex)) +
  geom_point() +
  scale_color_hue(
    name = "Sex", 
    labels = c("Females", "Males")) +
  xlab('Age Group (years)') +
  ylab('EEOP Rate (%)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
```


```{r}
# create a two-panel figure and show the plot

fhp_plot / eeop_plot
```

From upper panel, mean value equals to median in most of the age groups, indicating a non-skewed distributions in those data. For females in 30's, 50's, >60's group, data distributions were right-skewed. Also, the males had relatively larger FHP size than females. 

From lower panel, overall trend of EEOP rate decreased first and then increased. It had largest rate in 18-29 years group and lowest rate in 40's group. Also, EEOP rates in males were higher than that in females at all age groups.

```{r}
fhp_eop_plot = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_size_mm) %>% 
  filter(eop_size_mm != 0) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm, color = sex)) +
  geom_point() +
  geom_smooth(
    method = "lm", 
    color = 'black',
    se = FALSE) +
  facet_grid(
    sex ~ age_group, 
    labeller = as_labeller(c('0' = 'Female', '1' = 'Male', '2' = '18-29 years', 
                             '3' = "30's", '4' = "40's", '5' = "50's", '6+' = ">60's"))) +
  xlab('FHP Size (mm)') +
  ylab('EOP Size (mm)') +
  scale_color_hue(
    name = "Sex", 
    labels = c("Females", "Males"))

fhp_eop_plot
```

For male group, the EOP size and FHP size had slightly positive relationships in all age groups. However, for female group, the relationships were not identical among different age groups.


## 3. Reproduce Reported Results

```{r}
mtp_data %>% 
  mutate(
    age_group = case_when(
      age_group == '1' ~ 'less than 18',
      age_group == '2' ~ '18-30',
      age_group == '3' ~ '31-40',
      age_group == '4' ~ '41-50',
      age_group == '5' ~ '51-60',
      age_group == '6+' ~ 'larger than 60'
    )
  ) %>% 
  group_by(age_group) %>% 
  summarize(n = n()) %>% 
  knitr::kable(
    col.names = c('Age Group', 'Sample Size')
  )
```

* Sample sizes in each age group from available data were not consistent with authors stated.

```{r}
# overall mean

mean(pull(mtp_data, fhp_size_mm), na.rm = TRUE)

# mean and std by sex group

mtp_data %>% 
  mutate(
    sex = case_when(
      sex == '0' ~ 'Female',
      sex == '1' ~ 'Male'
    )
  ) %>%
  group_by(sex) %>% 
  summarize(
    maen = mean(fhp_size_mm, na.rm = TRUE), 
    sd = sd(fhp_size_mm, na.rm = TRUE)
  ) %>% 
  knitr::kable(
    col.names = c('Sex', 'Mean (FHP size)', 'Standard Deviation (FHP size)')
  ) 
```

* Reported mean and standard deviations for FHP size were consistent with that in available data.

```{r}
mtp_data %>% 
  mutate(
    eeop = case_when(
      eop_size_mm > 10 ~ 'EEOP',
      eop_size_mm <= 10 ~ 'Not EEOP'
    )
  ) %>% 
  count(eeop) %>%
  mutate(rate = prop.table(n)) %>% 
  knitr::kable(
    col.names = c('EEOP or not', 'Number', 'Prevalence')
  )
```

* Bony outgrowths were named as EEOP when they exceeded 10 mm in size. Variable used to evaluate the prevalence of EEOP was `eop_size_mm`. The EEOP prevalence from available data was 32%, which has 1% difference from what authors reported.

```{r}
fhp_violin = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_category) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm)) +
  geom_violin() +
  xlab('Age Group') +
  ylab('FHP Size (mm)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
  
fhp_col = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_category) %>% 
  group_by(age_group) %>% 
  count(fhp_category) %>% 
  mutate(rate = prop.table(n)) %>%  
  ggplot(aes(x = fhp_category, y = rate)) +
  geom_col() +
  xlab('FHP Categroy') +
  ylab('Percentage') +
  facet_grid(~age_group) 

fhp_violin / fhp_col

mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_category) %>% 
  group_by(age_group) %>% 
  count(fhp_category) %>% 
  mutate(rate = prop.table(n)) %>%
  filter(age_group == '6+' & fhp_category %in% c('4', '5', '6', '7')) %>% 
  summarize(sum = sum(rate)) %>% 
  knitr::kable(
    col.names = c('Age Group', 'Percentage of FHP >40 mm')
  )
```

* In older age group, more people had higher FHP size and the percentages of large-FHP categories also increased. Thus, the broad trends are consistent with that in the article. However, based on available data, the percentage that FHP >40 mm observed is 32.5%, which is not consistent with the article.  

## 4. Discussions

Based on the available data, lots of misclassification were found in category variables and mistakes were found in data entry, which made the data analyses in original report questionable. Some presentations of results also lacked important information. Besides, some statistical values were found not consistent with that reported in the article. Although board trends were consistent with the article that younger individuals had higher EEOP rates, we still cannot conclude cell phones cause hron growth. Extra data showing the prevalence of cell-phones usage among different age groups will be necessary to better address such hypothesis. 

