---
title: "P8105 Midterm Project"
author: "Junxian Chen (jc5314)"
date: "10/18/2019"
output: github_document
---

```{r, include = FALSE}
library(readxl)
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_gray(base_size = 10) + theme(legend.position = "bottom"))
```


## Introduction:

An article published on Nature Scientific Reports described relationship between forward head protraction (FHP) and enlarged external occipital protuberance (EEOP). The researchers analyzed data from individuals with different ages and sex, and pointed out that the prominent exostosis were more substantial and prevalent in young adults than older groups. However, people were skeptical about this study and a published author correction  indicated that errors or defects existed in the original work. In this project, the data that accompany the author correction are used to test whether mistakes are still exist and the results are consistent with that reported in the article.


## 1. Dataset

### 1.1 Dataset Description

```{r}
raw_data = 
  read_excel("./data/p8105_mtp_data.xlsx", skip = 8)
```

```{r}
mtp_data =   
  raw_data %>% 
  janitor::clean_names() %>% 
  replace_na(list(eop_size_mm = 0, eop_shape = 0)) %>% 
  mutate(
    sex = as.character(sex),
    age = as.integer(age),
    age_group = case_when(
      age_group %in% c('6', '7', '8') ~ '6+', 
      TRUE ~ age_group ),
    age_group = factor(age_group, levels = c('1', '2', '3', '4', '5', '6+')),
    eop_size = factor(eop_size, levels = c('0', '1', '2', '3', '4', '5')),
    eop_visibility_classification = factor(eop_visibility_classification, levels = c('0', '1', '2')),
    eop_shape = as.character(eop_shape),
    fhp_category = factor(fhp_category, levels = c('0', '1', '2', '3', '4', '5', '6', '7'))
  )

str(mtp_data)
```

The raw dataset was imported and cleaned. (Because the raw dataset would be used later to illustrate mistakes, the process of importing and cleaning were seperated here). During the data cleaning, missing values were filled according to the header information, variable types were modified, and orders were given to the factor variables. Since the article treated individuls older than 60 as '>60' group, those people were all assigned as '6+' under the `age_group` variables here to facilitate furthur analyses.  

After data cleaning, the resulting dataset, named `mtp_data`, contains `r nrow(mtp_data)` observations (i.e. the number of participants) and `r ncol(mtp_data)` variables. Key variables: The average value in **age** was `r mean(pull(mtp_data, age))`. The average **EOP size** was `r mean(pull(mtp_data, eop_size_mm))` mm. The average **FHP size** was `r mean(pull(mtp_data, fhp_size_mm), na.rm = TRUE)` mm. It was noticed that `NA` exist in the `fhp_size_mm` and `eop_shape`. Since no information was provided, the `NA`'s here were left blank.

A table below shows the age and gender distribution.

```{r}
mtp_data %>% 
  mutate(
    age_group = case_when(
      age_group == '1' ~ 'less than 18',
      age_group == '2' ~ '18-30',
      age_group == '3' ~ '31-40',
      age_group == '4' ~ '41-50',
      age_group == '5' ~ '51-60',
      age_group == '6+' ~ 'larger than 60'
    )
  ) %>% 
  group_by(sex, age_group) %>% 
  summarize(n = n()) %>%
  pivot_wider(
    names_from = sex,
    values_from = n
  ) %>% 
  knitr::kable(
    col.names = c('Age', 'Female', 'Male'),
  )
```

### 1.2 Issues in Dataset:

```{r}
mtp_data %>% 
  ggplot(aes(x = age_group, y = age, color = age_group)) +
  geom_point() +
  geom_hline(yintercept = c(18, 31, 41, 51, 61), color = 'red', linetype = 'dotted') + 
  xlab('Age Group') +
  ylab('Age (years)') +
  scale_color_hue(name = "age group") +
  scale_x_discrete(
    labels = c("<18's", "18-29 years","30's", "40's", "50's", ">60's")) +
  scale_color_hue(name = "age group",
                  labels = c("<18's", "18-29 years","30's", "40's", "50's", ">60's"))
```

* An outlier in the '<18's' group indicates a person older than 18 was mistakenly labled into this group. 

```{r}
mtp_data %>% 
  ggplot(aes(x = eop_size, y = eop_size_mm, color = eop_size)) +
  geom_point() +
  geom_hline(yintercept = c(5, 10, 15, 20, 25), color = 'red', linetype = 'dotted') + 
  xlab('EOP Size Group') +
  ylab('EOP Size (mm)') +
  scale_color_hue(name = "eop size group")
```

* Outliers indicate more misclassifications in EOP size group, especially in group 1, 3, 4, 5. The `NA` in the x-axis indicates an abnormal value in EOP size group.

```{r}
row_na = which(is.na(pull(mtp_data, eop_size)))
raw_data[row_na,] %>% 
  knitr::kable()
```

After tracing back, the induced `NA` is due to the value '14.6' in 'EOP size', which suppose to be 0 ~ 5, in raw dataset.

```{r}
mtp_data %>% 
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = eop_visibility_classification)) +
  geom_point() +
  geom_hline(yintercept = c(0, 5), color = 'red', linetype = 'dotted') + 
  xlab('EOP Visibility Classification') +
  ylab('EOP Size (mm)') +
  scale_color_hue(name = "EOP visibility classification")
```

* Outliers indicate misclassifications exist in all three groups in EOP visibility classification.

```{r}
mtp_data %>% 
  ggplot(aes(x = fhp_category, y = fhp_size_mm, color = fhp_category)) +
  geom_point() +
  geom_hline(yintercept = c(0, 10, 20, 30, 40, 50, 60, 70), color = 'red', linetype = 'dotted') + 
  xlab('FHP Size Group') +
  ylab('FHP Size (mm)') +
  scale_color_hue(name = "FHP size group")
```

* Outliers indicate misclassifications exist in FHP size group, especially in group 1, 2, 4. An `NA` in x-axis indicates an abnormal value in raw data. 

```{r}
row_na = which(is.na(pull(mtp_data, fhp_category)))
raw_data[row_na,] %>% 
  knitr::kable()
```

After tracing back, the induced `NA` is due to the '30.8' in 'FHP Category', which suppose to be 0 ~ 7, in raw dataset.


## 2. Visualization Improvement

```{r}
fhp_plot =
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot() +
  stat_summary(
    fun.y = mean, aes(colour = "Mean"), 
    geom = "point", 
    shape = 18, 
    size = 3,
    position = position_dodge(width = 0.75)) +
  stat_summary(
    fun.data = mean_se, 
    geom = "errorbar",
    width = 0.15,
    position = position_dodge(width = 0.75)) +
  scale_fill_hue(
    name = "Sex", 
    labels = c("Females", "Males")) +
  scale_colour_manual(
    values = c("Mean" = "darkred"), 
    labs(colour = '')) +
  xlab('Age Group (years)') +
  ylab('FHP Size (mm)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
```


```{r}
eeop_plot = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  mutate(
    eeop = case_when(
      eop_size_mm > 10 ~ 'y',
      eop_size_mm <= 10 ~ 'n'
    )
  ) %>% 
  group_by(age_group, sex) %>% 
  count(eeop) %>%
  mutate(rate = prop.table(n)) %>% 
  filter(eeop == 'y') %>% 
  ggplot(aes(y = rate, x = age_group)) +
  geom_line(aes(group = sex, color = sex)) +
  geom_point() +
  scale_color_hue(
    name = "Sex", 
    labels = c("Females", "Males")) +
  xlab('Age Group (years)') +
  ylab('EEOP Rate (%)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
```


```{r}
# create a two-panel figure and show the plot

fhp_plot / eeop_plot
```

In the upper panel, mean value equals to median in most of the age groups, indicating a non-skewed distributions. Also, the males had relatively larger FHP size than females. From the lower panel, overall trend of EEOP rate decreased first and then increased. It had largest rate in 18-29 years group and lowest rate in 40's group. Also, EEOP rate in males higher than that in females at all age groups.


```{r}
fhp_eop_plot = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_size_mm) %>% 
  filter(eop_size_mm != 0) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm, color = sex)) +
  geom_point() +
  geom_smooth(
    method = "lm", 
    color = 'black',
    se = FALSE) +
  facet_grid(
    sex ~ age_group, 
    labeller = as_labeller(c('0' = 'Female', '1' = 'Male', '2' = '18-29 years', 
                             '3' = "30's", '4' = "40's", '5' = "50's", '6+' = ">60's"))) +
  xlab('FHP Size (mm)') +
  ylab('EOP Size (mm)') +
  scale_color_hue(
    name = "Sex", 
    labels = c("Females", "Males"))

fhp_eop_plot
```

In male group, the EOP size and FHP size had slightly positive relationships in all age groups. However, in female group, the relationships were not indentical among age groups.


## 3. Reproducing reported results

```{r}
mtp_data %>% 
  mutate(
    age_group = case_when(
      age_group == '1' ~ 'less than 18',
      age_group == '2' ~ '18-30',
      age_group == '3' ~ '31-40',
      age_group == '4' ~ '41-50',
      age_group == '5' ~ '51-60',
      age_group == '6+' ~ 'larger than 60'
    )
  ) %>% 
  group_by(age_group) %>% 
  summarize(n = n()) %>% 
  knitr::kable(
    col.names = c('Age Group', 'Sample Size')
  )
```

* Sample sizes in each age group from available data were not consistent with authors' stated.

```{r}
# overall mean

mean(pull(mtp_data, fhp_size_mm), na.rm = TRUE)

# mean and std by sex group

mtp_data %>% 
  group_by(sex) %>% 
  summarize(
    maen = mean(fhp_size_mm, na.rm = TRUE), 
    sd = sd(fhp_size_mm, na.rm = TRUE)
  ) %>% 
  knitr::kable(
    col.names = c('Sex', 'Mean', 'Standard Deviation')
  ) 
```

* Reported mean and standard deviations for FHP size were consistent with that in avaliable data.

```{r}
mtp_data %>% 
  mutate(
    eeop = case_when(
      eop_size_mm > 10 ~ 'EEOP',
      eop_size_mm <= 10 ~ 'Not EEOP'
    )
  ) %>% 
  count(eeop) %>%
  mutate(rate = prop.table(n)) %>% 
  knitr::kable(
    col.names = c('EEOP or not', 'Number', 'Prevalence')
  )
```

* Those bony outgrowths were named as EEOP when they have exceeded 10 mm in size. Variable used to evalute the prevalence of EEOP was `eop_size_mm`. The EEOP prevalence from avaliable data is 32%, which has 1% difference from what authors reported.

```{r}
p1 = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_category) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm)) +
  geom_violin() +
  xlab('Age Group') +
  ylab('FHP Size (mm)') +
  scale_x_discrete(
    labels = c("18-29 years","30's", "40's", "50's", ">60's"))
  
p2 = 
  mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_category) %>% 
  group_by(age_group) %>% 
  count(fhp_category) %>% 
  mutate(rate = prop.table(n)) %>%  
  ggplot(aes(x = fhp_category, y = rate)) +
  geom_col() +
  xlab('FHP Categroy') +
  ylab('Percentage') +
  facet_grid(~age_group) 

p1 / p2

mtp_data %>% 
  filter(age_group != '1') %>% 
  drop_na(fhp_category) %>% 
  group_by(age_group) %>% 
  count(fhp_category) %>% 
  mutate(rate = prop.table(n)) %>%
  filter(age_group == '6+' & fhp_category %in% c('4', '5', '6', '7')) %>% 
  summarize(sum = sum(rate)) %>% 
  knitr::kable(
    col.names = c('Age Group', 'Percentage of FHP >40 mm')
  )
```

* From the two-panel polt, it can be seen that in older age group, more people had higher FHP size and percentages of large-FHP categroies increased in older age group. Thus, the broad trends are consistent with our avaliable data. However, based on avaliable data, the percentage that FHP >40 mm observed is 32.5%, which is not consistent with what reported in the article.  

## 4. Summary and Discussions

Summarize your results, the quality of the data analysis / presentation of results in the original report, and comment on the conclusions of the reports’ authors. Do you think the data provide evidence that cell phones are causing horn growth? What other data would you like to have to address this hypothesis?



